---
title: "Machine Learning in R using mlr on the AirBnb Dataset"
author: "Jakob Richter"
date: "November 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.width = 8)
```

```{r download data, include=FALSE}
if (!file.exists("data/listings.csv")) {
  url = 'http://data.insideairbnb.com/germany/be/berlin/2017-05-08/data/listings.csv.gz'
  download.file(url = url, destfile = "listings.csv.gz")  
  if (.Platform$OS.type == "unix") {
    system("gzip -d data/listings.csv.gz")
  } else {
    stop("Please extract the file yourself.")
  }
}
```

## The libraries we will use

```{r libraries}
library(data.table) # faster then data.frames
library(stringi) # fast string operations
library(ggplot2)
library(leaflet) # interactive maps
library(plyr) # revalue etc.
library(dplyr) # tidy data
library(magrittr) # pipe %>% etc.
# We need the latest version of mlr so you might want to run
# library(devtools)
# install_github("mlr-org/mlr")
library(mlr) # machine learning
library(gender) # gender for names
```

## Read the data

```{r read data}
data.listings = fread("data/listings.csv")
```

Convert strings like `"500$"` and `"30%"` to numbers.

```{r dollars}
col.with.dollars = c("price", "weekly_price", "monthly_price", "security_deposit", "cleaning_fee", "extra_people")
data.listings %<>% mutate_at(col.with.dollars, funs(as.numeric(stri_replace_first_fixed(., '$', ''))))
data.listings %<>% mutate_at("host_response_rate", funs(as.numeric(stri_replace_first_fixed(., '%', ''))))
data.listings = as.data.table(data.listings)
```

## Prepare for Machine Learning

Define which variables can be used to predict the *price* and which not.

```{r define vars}
learnable = c("latitude", "longitude", "property_type", "room_type", "accommodates", "bathrooms", "bedrooms", "beds", "host_is_superhost", "neighbourhood_cleansed", "cancellation_policy", "reviews_per_month", "calculated_host_listings_count", "require_guest_phone_verification", "require_guest_profile_picture", "instant_bookable", "number_of_reviews", "review_scores_value", "review_scores_location", "review_scores_communication", "review_scores_checkin", "review_scores_cleanliness", "review_scores_accuracy", "review_scores_rating", "availability_365", "minimum_nights", "maximum_nights", "square_feet", "bed_type", "host_response_rate", "host_response_time", "host_location", "guests_included")
```

Have we missed anything?
```{r check vars}
# typos?
setdiff(learnable, colnames(data.listings))

not.learnable = setdiff(colnames(data.listings), learnable)
glimpse(data.listings[, not.learnable, with = FALSE])
```

### Feature Generation

Add column for gender of host

```{r gender}
names.gender = gender(unique(data.listings$host_name))
names.gender = dplyr::rename(names.gender, host_name = name) #do not confuse with plyr::rename!!
data.listings = merge(data.listings, names.gender[,c("host_name", "proportion_female")])
learnable = c(learnable, "proportion_female")
```

### Inspect Data

How does the data look we want to learn on?

```{r view learnable}
glimpse(data.listings[, learnable, with = FALSE])

data = data.listings[reviews_per_month > 0.1 & !is.na(price), c(learnable, "price"), with = FALSE]
```

Prepare variables

```{r prepare vars}
# only differeciate between berlin and not berlin host.
data[, host_is_berlin := stri_detect_fixed(host_location, pattern = "Berlin")]
data$host_location = NULL

# convert host response time to ranking
data$host_response_time %<>% revalue(c("within an hour" = 1, "within a day" = 3, "N/A" = 5, "a few days or more" = 4, "within a few hours" = 2)) %>% as.integer()

# convert cancellation_policy to ranking
data$cancellation_policy %<>% revalue(c("strict" = 3, "flexible" = 1, "moderate" = 2, "super_strict_60" = 4)) %>% as.integer()


# convert columns with t/f as character to logical
tf.cols = names(which(sapply(data, function(x) all(c("t","f") %in% unique(x)))))
data %<>% mutate_at(tf.cols, funs('=='(.,"t")))

# convert strings to factors
char.cols = names(which(sapply(data, function(x) is.character(x) || is.logical(x))))
data %<>% mutate_at(char.cols, funs(as.factor))
```

## Machine Learning

Fit a **random forest** model on the data with the `price` as label.

```{r ml train}
tsk = makeRegrTask(data = data, target = "price")
tsk = removeConstantFeatures(tsk)

# impute of missing values
lrn = makeLearner("regr.ranger", num.trees = 200, importance = "permutation")
lrn = makeImputeWrapper(lrn, classes = list(integer = imputeMax(10), numeric = imputeMax(10), factor = imputeConstant("__miss__")))
tsk
system.time({
  mod = train(lrn, task = tsk)
})
```

Let's predict the prices using the trees that have not seen those observations.
This is important because otherwise we would have overfitted predictions.
We call that **out of bag predictions**.

```{r ml oob}
p = getOOBPreds(mod, tsk)
```

Let's see how well do we predict the prices and if we have outliers.
```{r ml vispred}
plot(p$data$truth, p$data$response, pch = 19, col = rgb(0,0,0,0.1))
abline(0,1, col = "red")
```

### Variable Importance

Which variables are the main predictors for the price?
```{r ml importance}
featimp = getFeatureImportance(mod)
plot.data = melt(featimp$res)
plot.data$variable %<>% reorder(plot.data$value)
g = ggplot(plot.data, aes(x = variable, y = value, fill = value))
g + geom_col() + coord_flip()
```

### Partial Dependency Plots

How do certain variables influence the price?
```{r ml partial}
pd = generatePartialDependenceData(mod, tsk, c("number_of_reviews", "accommodates", "host_response_time", "cancellation_policy"), individual = FALSE)
g = plotPartialDependence(pd)
g

pd = generatePartialDependenceData(mod, tsk, c("number_of_reviews", "accommodates", "host_response_time", "cancellation_policy"), individual = TRUE)
g = plotPartialDependence(pd, p = 0.02)
g
```

### Find those cheap flats

```{r ml cheap}
important.vars = c("latitude", "longitude", "property_type", "bathrooms", "bedrooms", "beds", "reviews_per_month", "number_of_reviews", "review_scores_value", "review_scores_location", "review_scores_communication", "review_scores_checkin", "review_scores_cleanliness", "review_scores_accuracy", "review_scores_rating", "square_feet", "bed_type", "host_response_rate", "host_response_time", "guests_included")
data = as.data.table(data)
data.sml = data[!is.na(price) & accommodates %in% 2:4 & room_type == "Entire home/apt" & beds %in% 1:2 & longitude > 13.28 & longitude < 13.52 & latitude > 52.45 & latitude < 52.57 , c(important.vars, "price"), with = FALSE]

tsk.sml = makeRegrTask(data = data.sml, target = "price")

mod.sml = train(lrn, task = tsk.sml)
p = getOOBPreds(mod.sml, tsk.sml)
plot(p$data$truth, p$data$response, pch = 19, col = rgb(0,0,0,0.1))
abline(0,1, col = "red")
```

We want to see where these values on the top left side are on the map:
```{r map}
data.sml$saving = p$data$response - p$data$truth

library(ggmap)
bbox = make_bbox(lon = data.sml$longitude, lat = data.sml$latitude)
map = get_map(location = bbox, source = "google")
g = ggmap(map, darken = c(0.3, "white"))
g = g + geom_point(data = data.sml[saving > 0], aes(x = longitude, y = latitude, color = saving, alpha = saving))
g = g + scale_alpha(range = c(0.01,1))
g = g + scale_color_continuous(low = "#333333", high = "#FF0000")
g
```

Using Leaflet to generate interactive map:
```{r leaflet}
getColor = function(x) {
  ifelse(x > 25, "green", ifelse(x < -25, "red", "orange"))
}

icons = awesomeIcons(
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(data.sml$saving)
)

m = leaflet(data = data.sml) %>% addTiles() %>% addAwesomeMarkers(~longitude, ~latitude, icon = icons, label = ~sprintf("Saving: %.2f, Price: %.2f", saving, price), clusterOptions = markerClusterOptions())
m
```

### Possible Improvements

* Put more weight to flats that are more often rented.
* Put more weight on recent bookings
* Compare price against actual bookings. Which flats get booked more often?
* Filter dubious ads.

## Resources

* mlr project page: https://github.com/mlr-org/mlr
* mlr tutorial: https://mlr-org.github.io/mlr-tutorial/devel/html/
* Vignette for `data.table`: https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html
* `leaflet` tutorial: https://rstudio.github.io/leaflet/
* This document was created using Rmarkdown and [`knitr`](https://yihui.name/knitr/)